%*******************************************************
% Publications
%*******************************************************
\pdfbookmark[1]{Publications}{Scientific Publications}
\chapter*{Scientific Publications}
Some of the ideas, images and data exposed in this thesis have been previously published in the next references:
\bigskip

\begin{itemize}

\item \person{Antonio Miguel Mora, Paloma de las Cuevas, Juan Julián Merelo Guervós} \emph{Going a Step Beyond the Black and White Lists for URL Accesses in the Enterprise by Means of Categorical Classifiers}.  IJCCI (ECTA) 2014: 125-134.

\item \person{Paloma de las Cuevas, Antonio Miguel Mora, Juan Julián Merelo Guervós, Pedro A. Castillo, Pablo García-Sánchez, Antonio Fernández-Ares} \emph{Corporate security solutions for BYOD: A novel user-centric and self-adaptive system.}.  Computer Communications, 68, 83-95 (2015).

\item \person{Paloma de las Cuevas, Juan Julián Merelo Guervós, Pablo García-Sánchez} \emph{Soft Computing Techniques Applied to Corporate and Personal Security.} In Proceedings of the Companion Publication of the 2015 Annual Conference on Genetic and Evolutionary Computation (pp. 1193-1196). ACM.

\item \person{Paloma de las Cuevas, Zaineb Chelly, Juan Julián Merelo Guervós, Anna I. Esparcia-Alcázar} \emph{An improved decision system for URL accesses based on a rough feature selection technique.} In Recent Advances in Computational Intelligence in Defense and Security (pp. 139-167). Springer International Publishing (2016).

\item \person{Paloma de las Cuevas, Pablo García-Sánchez, María I. García-Arenas, Juan Julián Merelo Guervós} \emph{Comparing individual representation and evaluation for automatic rule extraction using GP in a BYOD scenario.}.  Natural Computing, (submitted, under review).

\end{itemize}
